The effect of starvation on the dynamics of consumer populations

### 2016-07-17 Manuscript/Yeakel_Kempes_Redner-7-10-2016.tex
some more text

%Extinction to the left and the right
Higher rates of starvation results in a larger flux of the population to the hungry state, eliminating reproduction and increasing the likelihood of death, however it is the relative size of the rate of starvation to the rate of recovery that determines the long-term dynamics of the system.
By looking at the ratio $\sigma/\rho$, we can assess the competing effects of cyclic dynamics and decreasing rates of extinction

>>>>

%Extinction to the left and the right
Higher rates of starvation results in a larger flux of the population to the hungry state, eliminating reproduction and increasing the likelihood of death, however it is the relative size of the rate of starvation to the rate of recovery that determines the long-term dynamics of the system.
By examining the ratio $\sigma/\rho$, we can understand the competing effects of cyclic dynamics and population density on the probability of extinction.
We computed the probability of extinction, which we defined as $H(t)+F(t)=10$ at any instant across all values of $t$, for 1000 replicates of the continuous-time system shown in Eq. XX with random starting conditions within the interval $(R^*,H^*,F^*)\pm10\%$ (check) for an organism of $M=100{\rm g}$.
By allowing the rate of starvation to vary, we assessed extinction risk across a range of the ratio $\sigma/\rho$ varying between XXX and XXX, thus examining a vertical cross-ection of Fig \ref{Hopfb}.
As we expected, we found higher rates of extinction for both low and high values of $\sigma/\rho$; for low values the higher extinction risk is the consequence of transient cycles with larger amplitudes as the system nears the Hopf bifurcation.
For high values of $\sigma/\rho$, higher extinction risk is due to the steady decrease in the steady state consumer population density.
This interplay creates an `extinction refuge', such that for a relatively small range of $\sigma/\rho$ extinction probabilities are minimized.

As has been described, the $\sigma$ vs. $\rho$ space cannot be freely traversed, such that not all values of $\sigma/\rho$ are biologically feasible.
We observe that the allometrically constrained values of $\sigma/\rho$ (with $\pm 20\%$ variability around the expectation) fall squarely within the extinction refugia, such that the are close enough to the Hopf bifurcation to avoid low steady state densities, but far enough away to avoid large-amplitude transient cycles.
The fact that allometric values of $\sigma$ and $\rho$ fall within this relatively small window is suggestive of a selective mechanism driving these rates within populations, and suggests that the NSM system provides insight to a general phenomena that may both drive -- and constrain -- natural animal populations.
